---
title: "Project-2"
output:
  word_document: default
  pdf_document:
    number_sections: yes
    latex_engine: xelatex
fig_caption: yes
---

```{r setup, include=FALSE, eval = TRUE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, warning=FALSE, message=FALSE)
```


```{r,echo = TRUE,message=FALSE}
library(dplyr)
library(moderndive)
library(ISLR)
library(skimr)
library(plotly)
library(tidyr)
library(jtools)
library(kableExtra)
#install.packages("corrplot")
library(corrplot)
library(ggplot2)
library(janitor)
```

# Introduction

The number of household members is one of the properties of household structure.Household unit is composed of family units. 
Sharma (2013) regarded the family as a functional unit of society. It functions in an economic, social or emotional sense.Hence, the structure of it and those influential factors behind it are of much importance. 

(Sharma R. (2013). The Family and Family Structure Classification Redefined for the Current Times. Journal of family medicine and primary care, 2(4), 306–310. https://doi.org/10.4103/2249-4863.123774)


# Exploratory analysis

```{r,}

# Dataset prepare
data <- read.csv("dataset3.csv")

# Ignore the region and rename
household <- data[,-2]
colnames(household) <- c("Income", "Food.exp", "Head.sex", "Head.age", "Type", "Members.no", "Area", 
                         "House.age", "Bedrooms.no", "Elec.")

# Check is there any NA in data
cat("\n Check is there any NA in data \n")
apply(household, 2, function(x) any(is.na(x))) # No NAs in data

# Move our response variable to the first
household <- household %>%
  relocate("Members.no") # Use dplyr::relocate to move to the front
head(household) %>%   
  kable(digits = 2) %>%
  kable_styling(latex_options = 'HOLD_position')

```
First, we look at the proportion of each no. of households.

```{r, fig.width = 6, fig.height = 4, fig.align = "center", fig.pos = "H"}
household <- household %>% 
  mutate(Members.group = as.character(Members.no))
household[household$Members.no == 1, "Members.group"] = c("1 person")
household[household$Members.no == 2, "Members.group"] = c("2 people") 
household[household$Members.no == 3, "Members.group"] = c("3 people") 
household[household$Members.no == 4, "Members.group"] = c("4 people") 
household[household$Members.no == 5, "Members.group"] = c("5 people") 
household[household$Members.no == 6, "Members.group"] = c("6 people") 
household[household$Members.no >= 7 & household$Members.no <=16, "Members.group"] = c("7 or more people") 

prop.y.group <- household %>%
  group_by(Members.group) %>%
  count()%>%
  mutate(prop = round(n / nrow(household), 2))
prop.y.group

ggplot(prop.y.group, aes(x = "", y = n, fill=Members.group)) +
         # Create a bar first
         geom_bar(width = 1, stat = "identity") +
          
         # Then do pie chart
         coord_polar("y", start=0) +
         # Use color palettes from RColorBrewer package
         scale_fill_brewer(palette="Dark2") +
         # Blank the axis text and labs
         theme(axis.text.x=element_blank()) +
         labs(x="", y="", title="Percentage of households by household size")

```

Assume the data had been random sampled, we can get a proportion of 49% of the households contained 5 or more people in this region of Philippine. 



```{r, fig.width = 6, fig.height = 4, fig.align = "center", fig.pos = "H"}
# summary of response varibale
household %>% select(Members.no)%>%
  skim()
unique(household$Members.no)

max.no <- max(household$Members.no)
ggplot(household, aes(x = Members.no)) +
  # Set bins as max(Members.no)
  geom_histogram(bins = max.no, fill = "#0038A8", col = "white") +
  labs(main = "Total number of family members in Northern Mindanao")

# We first cut Y into classes
median.no <- quantile(household$Members.no, probs = c(0.5));median.no
# Median.no is 4

# Binary classification
points1 <- c(1, 4, 16)

household$quantiles <- cut(household$Members.no, breaks = points1,
                           include.lowest = TRUE, labels = c("1-4", "5-16"))
# household$quantiles <- cut(household$Members.no, breaks = points2,
#                           include.lowest = TRUE, labels = c("1-4","5-8","9-12","13-16"))

prop.y.group <- household %>%
  group_by(quantiles) %>%
  count()
  
ggplot(prop.y.group, aes(x = "", y = n, fill=quantiles)) +
         # Create a bar first
         geom_bar(width = 1, stat = "identity") +
         # Then do pie chart
         coord_polar("y", start=0) +
         # Use color palettes from RColorBrewer package
         scale_fill_brewer(palette="Dark2") +
         # Blank the axis text and labs
         theme(axis.text.x=element_blank()) +
         labs(x="", y="", title="Percentage of households by household size")

```



## Skim of xs

```{r,}
household <- household %>%
  mutate(Income.kPHP = Income/1000, Food.exp.kPHP= Food.exp/1000)

group.vars <- c("Elec.", "Type", "Head.sex")
num.vars <- c("Income.kPHP", "Food.exp.kPHP", "Area", "House.age", "Bedrooms.no", "Head.age")
# Code from DAS example sourcecode
func_my_skim <- skim_with()
func_my_skim(household[,num.vars])  %>%
  dplyr::select(-skim_type) %>%
  kable(col.names = c("Variable", "Missing", "Complete", "Mean", "SD", "Min.", "1st Q.", "Median",
                        "3rd Q.", "Max.", "Hist"),
        caption = '\\label{tab:summary} Summary statistics on teaching and beauty scores.',
        booktabs = TRUE, digits = 2) %>%
  kable_styling(font_size = 10, latex_options = "hold_position")

```
The economic numeric variables contain much variability. Considering the property of linear model, a log-transformation is suitable for those economic covariates.

### Check Y vs. numeric variables

```{r, }
## Why we use log-transform

# Plotting with continuous variables
# Income, Food.exp, head.age
# ------------------ Income------------------------- #
p1 <- ggplot(data = household, aes(x = as.factor(quantiles), y = Income, fill = as.factor(quantiles))) +
    geom_boxplot() +
    labs(x = "", y = "Income")+ 
    theme(legend.position = "none", title = element_text(size = 16)) +
    labs(title = "Members number by Income")

p1.log <-ggplot(data = household, aes(x = as.factor(quantiles), y = log(Income), fill = as.factor(quantiles))) +
    geom_boxplot() +
    labs(x = "", y = "Log Income")+ 
    theme(legend.position = "none", title = element_text(size = 16)) +
    labs(title = "Members number by log Income")
p1; p1.log
```


We applied log-transform for *Income*, to stable the variability.
Similarly, we did to *Food.exp*, *Area*


```{r, fig.width = 9, fig.height = 6, fig.align = "center", fig.pos = "H"}
library(gridExtra)

# Plotting with numerical variables
# Income, Food.exp, head.age
# ------------------ Income------------------------- #
p1 <-ggplot(data = household, aes(x = as.factor(quantiles), y = log(Income), fill = as.factor(quantiles))) +
    geom_boxplot() +
    labs(x = "", y = "Log Income")+ 
    geom_hline(yintercept = median(log(household$Income)),
                              linetype = "dashed") +
    theme(legend.position = "none") +
    labs(title = "Members number by log income")

# ------------------ Food.exp ------------------------- #
p2 <- ggplot(data = household, aes(x = as.factor(quantiles), y = log(Food.exp), fill = as.factor(quantiles))) +
    geom_boxplot() +
    labs(x = "", y = "Log food.exp")+ 
    geom_hline(yintercept = median(log(household$Food.exp)),
                              linetype = "dashed") +
    theme(legend.position = "none") +
    labs(title = "Members number by log food.exp")


grid.arrange(p1, p2,  ncol = 2)
```
The larger size households tend to earn more income and spend more on food expenditure.

In food expenditure, the difference gap is wider, since the 25% percentile of larger households’ spending nearly reach the 75% percentile of the smaller ones.



```{r, fig.width = 9, fig.height = 9, fig.align = "center", fig.pos = "H"}
library(gridExtra)

# ------------------ House.age ------------------------- #
p3 <- ggplot(data = household, aes(x = as.factor(quantiles), y = House.age, fill = as.factor(quantiles))) +
    geom_boxplot() +
    labs(x = "", y = "house.age")+ 
    theme(legend.position = "none") +
    geom_hline(yintercept = median(household$House.age), 
               linetype = "dashed") +
    labs(title = "Members number by house age")

# ------------------ Head.age ------------------------- #
p4 <- ggplot(data = household, aes(x = as.factor(quantiles), y = Head.age, fill = as.factor(quantiles))) +
    geom_boxplot() +
    labs(x = "", y = "Head.age")+ 
    # Add a mean age line
    geom_hline(yintercept = median(household$Head.age), 
               linetype = "dashed") +
    theme(legend.position = "none") +
    labs(title = "Members number by head age")

# ------------------ Bedrooms.no ------------------------- #
p5 <- ggplot(data = household, aes(x = as.factor(quantiles), y = Bedrooms.no, fill = as.factor(quantiles))) +
    geom_boxplot() +
    labs(x = "", y = "Bedrooms.no")+ 
    theme(legend.position = "none") +
    labs(title = "Members number by bedroom number")

# ------------------ Area ------------------------- #
p6.log <-ggplot(data = household, aes(x = as.factor(quantiles), y = log(Area), fill = as.factor(quantiles))) +
    geom_boxplot() +
    labs(x = "", y = "Log Area")+ 
    theme(legend.position = "none") +
    labs(title = "Members number by log area")

# p3, p4 are significantly different
# While other two are not
grid.arrange(p3, p4, p5, p6.log, ncol = 2)

```

Short conclusion:

Similarly, we checked other numeric variables, and two of them (house bedroom number and floor area) are found not significantly different between the two groups.

And the slight difference in the house age and head age may raise concern. We should check whether it is a significant predictor of the odds of a household being small or large size.

## Visualization of correlation

```{r,  fig.width = 9, fig.height = 6, fig.align = "center", fig.pos = "H"}


household <- mutate(household, 
                    Income.log = log(Income),
                    Food.exp.log = log(Food.exp))

#Corrplot
household %>%
    select(Members.no, Income.log, Food.exp.log, Head.age, House.age) %>%
    cor() %>%
    # Code reference:
    # http://www.sthda.com/english/wiki/visualize-correlation-matrix-using-correlogram
    corrplot(type="upper",
             addCoef.col = "black", # Add coefficient of correlation
             tl.col="black", tl.srt=45, #Text label color and rotation
             diag=FALSE) # hide correlation coefficient on the principal diagonal
```
There are high correlations between annual income and annual food expenditure.

The economic issues are among the top points in the government concerning. Find out the method to include these two core economic influences!

## Income groups

```{r, fig.width = 8, fig.height = 6, fig.align = "center", fig.pos = "H"}
library(gridExtra)

prop.points <- quantile(log(household$Income), probs=c(0, 0.25, 0.5, 0.75, 1.0))
# Back the log transfrom
income.points <- exp(prop.points);income.points
##      0%        25%        50%        75%       100% 
##  16238.00   85544.49  131806.00  249175.42 2598050.00 

household <- household %>%
  # Add Income groups
  mutate(income.groups = cut(Income/1000, breaks=c(0, 132, 249, 1e10),labels=c("< 132 thousand PHP", "132 - 249", "249 and more")))

p2 <-  ggplot(household, aes(x = log(Food.exp), y = Members.no)) +
    geom_point(aes(color = income.groups , fill = income.groups ), size=2, alpha = 0.4) +
    geom_smooth(method = "lm", se = FALSE,aes(color = income.groups , fill = income.groups ), size = 1)+
    xlim(9, 13) + ylim(0, 10.5) +
    theme(legend.position = "bottom")
  
p2
```


## Check group variables

```{r,}
# Code from lab9
# use the janitor package to summarise this data in a table format:
household %>% 
  tabyl(Head.sex, quantiles) %>% 
  adorn_percentages() %>% 
  adorn_pct_formatting() %>% 
  adorn_ns() # To show original counts

p1 <- ggplot(household, aes(x= as.factor(quantiles),  y = ..prop.., group=Head.sex, fill=Head.sex)) + 
    geom_bar(position="dodge", stat="count") +
    labs(y = "Proportion",x = "", main = "Barplot of Members.no by Head.sex")

# Check Type
household %>% 
  tabyl(Type, quantiles) %>% 
  adorn_percentages() %>% 
  adorn_pct_formatting() %>% 
  adorn_ns() # To show original counts

p2 <- ggplot(household, aes(x= as.factor(quantiles),  y = ..prop.., group=Type, fill=Type)) + 
    geom_bar(position="dodge", stat="count") +
    labs(y = "Proportion",x = "", main = "Barplot of Members.no by type")

# Check Electricity
household %>% 
  tabyl(Elec., quantiles) %>% 
  adorn_percentages() %>% 
  adorn_pct_formatting() %>% 
  adorn_ns() # To show original counts
household$Elec. <- as.factor(household$Elec.)
p3 <- ggplot(household, aes(x= as.factor(quantiles),  y = ..prop.., 
                            group=Elec., fill=Elec.)) + 
    geom_bar(position="dodge", stat="count") +
    labs(y = "Proportion",x = "", main = "Barplot of Members.no by electricity") +
    ylim(0, 0.6)
p3

library(gridExtra)
p1 <- p1 + theme(legend.position = "bottom")
p2 <- p2 + theme(legend.position = "bottom")
p3 <- p3 + theme(legend.position = "bottom")
grid.arrange(p1, p2, p3,  ncol = 2)

```


Short conclusion:

- For *Head.sex*, a large proportion in small households are female leaders (67.8%), while in more-member households are witnessed with more male heads (53.7%).

- For *Electricity*, the difference is not obvious with fewer-member households may have a percentage of 50.6% having access to electricity, while 49.0% for larger household. However, the two groups still have different electricity means.

- For *Type*, there are some interesting findings.

```{r, fig.width = 9, fig.height = 6, fig.align = "center", fig.pos = "H"}

type1 <- ggplot(data = household, aes(x = Type, y = Members.no, fill = Type)) +
    geom_boxplot() +
    labs(x = "Type", y = "Members.no") + 
    theme(legend.position = "none", title = element_text(size = 16))

household <- household %>% 
  mutate(Type.single = ifelse(Type=="Single Family", "Yes", "No"))
type2 <- ggplot(data = household, aes(x = Type.single, y = Members.no, fill = Type.single)) +
    geom_boxplot() +
    labs(x = "Type single", y = "Members.no") + 
    theme(legend.position = "none", title = element_text(size = 16))

grid.arrange(type1, type2,ncol = 2)

```



# Modelling
Note by Huan: Maybe Poisson model is more appropriate? The meeting can be discussed together~~
The modelling details are all there, confidence intervals and predictions are done, thanks for xuan working hard!

```{r,echo = FALSE}
library("janitor")

data_cut <- household[order(household$Members.no, decreasing = T),]
data_cut <- data_cut %>%
    mutate(Num2 = data$Total.Number.of.Family.members)
data_cut[,"Num2"] <- ifelse(data_cut[,"Num2"] > median(data_cut[,"Num2"]),"High","Low")

data_sex2 <- data_cut %>%
    select(Num2, Household.Head.Sex)

#look
data_sex2 %>% 
  tabyl(Household.Head.Sex, Num2) %>% 
  adorn_percentages() %>% 
  adorn_pct_formatting() %>% 
  adorn_ns()

ggplot(data_sex2, aes(x= Num2,  y = ..prop.., group=Household.Head.Sex, fill=Household.Head.Sex)) + 
    geom_bar(position="dodge", stat="count") +
    labs(y = "Proportion")

model_sex2 <- glm(Num2 ~ Household.Head.Sex, data = data_sex2, family = binomial(link = "logit"))
model_sex2 %>%
  summary()
levels(data_sex2$Household.Head.Sex)
confint(model_sex2) %>%
  kable()

mod.ethnic.coef.logodds <- model_sex2 %>%
                            summary() %>%
                            coef()

mod.ethnic.coef.logodds <- model_sex2 %>%
                            summary() %>%
                            coef()

mod.ethnic.coef.logodds <- model_sex2 %>%
                            summary() %>%
                            coef()

plot_model(model_sex2, show.values = TRUE, transform = NULL,
           title = "Log-Odds (Male instructor)", show.p = FALSE)
```

#Cut 5
```{r, echo = FALSE}
data_cut <- read.csv("C:/Users/Seismic/Downloads/dataset3.csv")
data_cut <- data_cut %>%
    mutate(Num5 = data$Total.Number.of.Family.members)
data_cut<-data_cut[order(data_cut$Num5,decreasing = T),]
data_cut$Num5<-cut(data_cut$Num5,breaks=c(1,3,6,10,16),labels=c('low1','low2','high1','high2'))



model_sex3 <-  multinom(Num5 ~ Household.Head.Sex, data = data_cut, probabilities = TRUE, model = TRUE)
summary(model_sex3)
Anova(model_type)
round(odds.ratio(model_type),2)

```

Delete income

```{r,}
data1 <- read.csv("dataset3.csv")
data1[,"Total.Number.of.Family.members"] <- ifelse(data1[,"Total.Number.of.Family.members"] > median(data1[,"Total.Number.of.Family.members"]),1,0)
data1$Total.Number.of.Family.members<-as.factor(data1$Total.Number.of.Family.members)

data1 <- data1 %>% 
  mutate(Type.single = ifelse(Type.of.Household =="Single Family", "Yes", "No"))

model3 <-  glm(Total.Number.of.Family.members ~ 
                   log(Total.Food.Expenditure)+
                   Household.Head.Age+
                   House.Age+
                   Electricity+
                   Household.Head.Sex+
                   Type.single ,family = binomial(link = "logit"),data = data1)
summary(model3)
```
```{r, }
# -------------------------------
model.full <- glm(Total.Number.of.Family.members ~ 
                   log(Total.Food.Expenditure)+
                   Household.Head.Age+
                   House.Age+
                   Electricity+
                   Household.Head.Sex+
                   Type.single ,family = binomial(link = "logit"),data = data1)
# --------------drop1------------------
drop1(model.full, test='Chisq') ## only for full model
# --------------lrtest-----------------
# lrtest(model.full, model.reduced)
## Interpretation - Pr(>Chisq) for reduced model is little, thus not accept the reduced model

```

# The comparisons
## Model4 + interactions

```{r,}
data1 <- data1 %>% 
  mutate(Head.Age.mean = Household.Head.Age - mean(Household.Head.Age))


ggplot(data =  data1, aes(x = as.factor(Total.Number.of.Family.members), y = Head.Age.mean, 
                          fill = as.factor(Total.Number.of.Family.members))) +
                geom_boxplot() +
                labs(x = "", y = "Head.Age.mean")+ 
                theme(legend.position = "none", title = element_text(size = 16)) +
                labs(title = "Members number by Head.Age.mean")

model4 <-  glm(Total.Number.of.Family.members ~ 
                   log(Total.Food.Expenditure) +
                   Head.Age.mean +
                   House.Age +
                   Electricity+
                   Household.Head.Sex +
                   Type.single ,family = binomial(link = "logit"),data = data1)
summary(model4)
```

# Poisson model

```{r,}
data1 <- read.csv("dataset3.csv")
data1 <- data1 %>% 
  mutate(Type.single = ifelse(Type.of.Household =="Single Family", "Yes", "No"))

poisson.model <- glm(Total.Number.of.Family.members ~ 
                       log(Total.Food.Expenditure) + 
                       Household.Head.Age + 
                       House.Age + 
                       Electricity + 
                       Household.Head.Sex + 
                       Type.single , family = poisson, data = data1)
summary(poisson.model)
```
## Residuals of Poisson【APM course】
```{r,fig.width = 9, fig.height = 9, fig.align = "center", fig.pos = "H"}

# Calculate the dispersion parameter:
pearson <- residuals(poisson.model, type = "pearson")
sum(pearson**2) / poisson.model$df.residual
## [1] 0.7135002

# Residual plots vs. predicted (using standardised residuals):
#------------------------------------------------
# Make use of rstandard()
# Get Std. residuals
stand.resid <- rstandard(model = poisson.model, type = "pearson")
fitted.values <-  predict(poisson.model, type = "response")# "response" = y, default = log(y)

plot(x = fitted.values, y = stand.resid, 
     xlab = "fitted.values", ylab = "Standardized Pearson residuals", ylim = c(-5,5))
abline(h = c(-3, -2, 0, 2, 3), lty = "dotted", col = "red")

```
